On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-06-09 reports/confidential/engineering.tex
mark white paper as confidential

\Paragraph{System Requirements}
In our running example, there is a system (a CPS) composed by $5$ main
subsystem which we consider atomic (i.e. not themself composed by subsystems)
and then agents: the tank, the sensor, the controller, the actuator, and the
motorized valve (valve from now on).  The whole specification (which we now
describe) of the physical and functional architecture is depicted in
Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
\begin{enumerate}
	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings. We assume a unidirectional communication without loss of generality}
	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
	\item the controller calculates the state of the valve and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
	\item the actuator translates (digital to analog) the state received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
\end{enumerate}
Each agent has an input or output port from which it communicates with other
agents.  It is also composed by the functional blocks that defines its
functional architecture and physical boundaries that defines its physical
architecture, as detailed in Appendix~\ref{app:requirements}.

@@

\Paragraph{Port Weaknesses}
From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item drop ($s\selectivedropport a$), 
		\item insert ($s\selectiveinsertport a$), or
		\item injection ($s\selectiveinjectionport a$)
\end{enumerate}
\end{enumerate}

\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
\begin{proof}
An input port is, in the $\abftheory$-theory, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base.  If
	a communication occurred and resulted in
	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
	$|\assertionRegion|\neq|\beliefRegion|$. 
	\begin{itemize}
		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
			a communication through the input-port), the two
			regions have a different number of subregions, and,
			then, there exist only two mutually exclusive options: 
			\begin{itemize}
				\item either $|\assertionRegion|>|\beliefRegion|$, where one
					ore more Assertions didn't become Beliefs, which we
					call \emph{drop port} ($s\dropport a$) since its behavior drops and
					prevents incoming communications,
				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
					one ore more Beliefs which have not been sent as
					input-Assertions, which we call \emph{insertion
					port} ($s\insertport a$) since its behavior generates new Beliefs
					unrelated to the incoming communication. 
			\end{itemize}
		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
				there exist only two mutually exclusive options: 
			\begin{itemize}
				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
					$\assertionRegion$ and $\beliefRegion$, i.e.
					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
					$t^0$ and $t^1$ represent time at the beginning of
					communication and right after, respectively); which
					means that Assertions have reached the port and new
					Beliefs have been generated by the port, but no
					correlation between the elements of the two regions exist. 
					We call this \emph{injection port} ($s\injectionport a$) since it has modified
					the information carried by the Assertions into new
					corresponding Beliefs.
				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
					this implies that no communication occurred, while we
					assumed that a communication was taking place (absurd).
			\end{itemize}
	\end{itemize}
	In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
	Similarly, 
	\begin{itemize}
		\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
			$\dr{\assertionRegion}{\beliefRegion}$ except for the
			overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
			$\eq{\assertionRegion}{\beliefRegion}$ (see
			\autocite{Santaca2016abf}),
			we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
		\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
			$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
			$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
		\begin{itemize}
			\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
				some Beliefs are not correlated to incoming Assertions
				and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
			\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
				some incoming Assertions are not correlated to Belief 
				and we call it \emph{selective drop port} ($s\selectivedropport a$)
		\end{itemize}
	\end{itemize}
\end{proof}

\subsubsection{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for the
sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

@@

\Paragraph{Channel Weaknesses}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any Assertion from its input-port to its output-port, there is no
insecure behavior but only the combination of the weaknesses of the input and
output port; therefore there exists $(7^2)-2=47$ theoretical configurations
($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on both input and oupput
side; and we exclude the configuration with $2$ secure types as input
and ouput, $-2$); where only 43 are possible.

\begin{enumerate}[start=5, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{secure output port} and \emph{input injection port} ($s\sjcom r$), 
	\item \emph{secure output port} and \emph{input selective drop port} ($s\ssdcom r$), 
	\item \emph{secure output port} and \emph{input selective insertion port} ($s\ssicom r$), 
	\item \emph{secure output port} and \emph{input selective injection port} ($s\ssjcom r$),

@@

As depicted in Figure~\ref{fig:atom}, so far, we have considered Information
generated by a Port $P_I$ and then sent through a channel $C$ to another
(recipient) port $P_O$. In this scenario where Ports and then Channels are
atomic (otherwise raising infinite regress), we can only  consider the
relations between Ports and Channel; considering both input-port to channel and
channel to output-port. In fact, the Weaknesses of a channel are defined in terms
of the Weaknesses of Ports. We now introduce the concept of
functional architecture and its decomposition into functional blocks. 
Io order to define a functional block without encountering a infinitely recursive definition,
we must reach the same conclusions as for the Channel. Therefore, describing
the information as flowing over a channel or in a functional block is purely
syntactic sugar.

@@

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent 
	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
	A functional architecture is an interconnected system of functional blocks.
\end{definition}
It is evident that the quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels). The behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).
Therefore, we can consider the relation between 
the outputs of a functional block with respect to the requirements.
One may, of course, argue that since there is a
transformation process in the functional block the quality of that
transformation process is what the theory should consider. This can
be argued not just for functional blocks (Behaviors) but also for Channels
(Assertions). The verification of the semantics of functional blocks and of the 
protocols executed over the channels will be considered at design level while
now we are still defining the specification.

\begin{itemize}
	\item $\po{\behaviorRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
	        output given the correct inputs. Not all the inputs are handled
	        properly, nor all the expected outputs are always generated
	        when correct inputs are given.
	\item $\pp{\behaviorRegion}{\factRegion}$ part of the expected outputs are not
	        generated in response to the correct
	        inputs
	\item $\ppi{\behaviorRegion}{\factRegion}$ the components
	        correctly performs the expected behavior when the correct
	        inputs are provided but is subject to input
	        injections
	\item $\dr{\behaviorRegion}{\factRegion}$ the component
		never performs the expected behavior (e.g. physical
		damage)
\end{itemize}
These weaknesses will be detailed afterwards in this section.

As we detail afterwards, the relations considered so
far revolves around the concept of integrity or availability, while the reminder of this section 
will be focused on a broader analysis of security properties and requirements;
which will also consider other security properties, such as availability and
confidentiality.  In order to define the quality of the internals of a functional block or channel, we first
need to relate the concept of Facts to engineering requirements.

\Paragraph{Requirements as Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there may exist a factorial requirement (Fact) predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

\Paragraph{Methodology}
We have now mapped the building blocks of our theory, namely Information,
Belief, Assertion, Behavior, and Fact. One may notice that the concept of
Knowledge hasn't been discussed in this section so far. Our argument is that
knowledge, as an epistemological concept cannot be defined due to several
implication that this would have (see, for example,
\autocite{Empiricus1990Pyrrhonism}). On the other hand, the factorial way in
which requirements are imposed may be true only in the specification, while not
necessarily true in the final implementation. While Knowledge may not be
apprehensible in general, we can define tests to see if Facts holds in a
design or implementation. What is of our interest is to guarantee security in a
system so we now correlate the potential behaviors with the standard (de-facto)
understanding of security; so that we can identify test that, at the end of the
engineering process, may not give Knowledge but, at last, tested Fact.

@@

\Paragraph{Availability and Integrity (Architecture Security)}
\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{client-server.pdf}
	\caption{Client-Server paradigm in $\abftheory$-theory}
	\label{fig:client-server}
\end{figure}
Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
in order to prevent a resource to be available, there
exists only (i.e. inclusively) the following cases
that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
\begin{enumerate}
	\item\label{case:srv1} The resource on the server cannot be found, so that it is
		possible for the client to contact the server and for the server
		to send a response back, but the output-port does not receive the
		necessary information. This may be due to a software error such
		that the functional architecture doesn't output a response, or
		a problem in the flow of information (i.e. the functional
		architecture doesn't receive the necessary information from
		other resources), or an hardware failure \&c. This can be formalized
		as an issue to be found in the relation between the beliefs produced by the functional
		architecture and the ones received by the output port, i.e.
		$\rcc(\behaviorRegion_s,\{\text{res}_0\})$ or 
		$\rcc(\beliefRegion_s,\{\text{res}_0\})$ 
	\item\label{case:srv2} The server hosting the resource is down, and then the server
		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
		requests (insecure input port due to drop-related issue) or both.
	\item\label{case:srv3} The client cannot properly compute request (symmetric to Case~\ref{case:srv1})
	\item The client is down, symmetric to Case~\ref{case:srv2}
	\item The channel is down due to a drop related issue in one of the two channels 
\end{enumerate}
We shall, then, distinguish between two cases, as we mentioned beforehand:
agent (or system) and Information (which we call Infosec) availability. 
\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
when the property of availability is expressed over an agent or a system, the
	list of possible designs do not include those in which there exist:
	drop ports, or drop channels/functional-blocks.
\end{definition}
\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
when the property of availability is expressed
	over a formula, the list of possible designs that violates the
	availability property is reduced to the part of the system that
	does not influence that specific formula.
\end{definition}

>>>>

\paragraph{System Requirements}
In our running example, there is a system (a CPS) composed by $5$ main
subsystem which we consider atomic (i.e. not themself composed by subsystems)
and then agents: the tank, the sensor, the controller, the actuator, and the
motorized valve (valve from now on).  The whole specification (which we now
describe) of the physical and functional architecture is depicted in
Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
\begin{enumerate}
	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings. We assume a unidirectional communication without loss of generality}
	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
	\item the controller calculates the state of the valve and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
	\item the actuator translates (digital to analog) the state received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
\end{enumerate}
Each agent has an input or output port from which it communicates with other
agents.  It is also composed by the functional blocks that defines its
functional architecture and physical boundaries that defines its physical
architecture, as detailed in Appendix~\ref{app:requirements}.

@@

\paragraph{Port Weaknesses}
From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item drop ($s\selectivedropport a$), 
		\item drop+insert ($s\selectiveinsertport a$), or
		%\item injection ($s\selectiveinjectionport a$)
\end{enumerate}
\end{enumerate}

\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
\begin{proof}
An input port is, in the $\abftheory$-theory, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between Regions that contain Information and Regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either $\assertionRegion=\beliefRegion=\emptyset$ (no communication) or $\assertionRegion=\beliefRegion\neq\emptyset$ (forward communication)
		\item if $\dr{\assertionRegion}{\beliefRegion})$ then $\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$ (we call \emph{insert} the former and \emph{full drop} the latter case), or $\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$ which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion})$ then $\beliefRegion$ contains and extend $\assertionRegion$ which we call \emph{injection}
		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then $\assertionRegion$ contains and extend $\beliefRegion$ which we call \emph{selective drop}
		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of the $\assertionRegion$ is contained in the $\beliefRegion$ which is a combination of \emph{selective drop and generation}
	\end{enumerate}

	%	If
	%	a communication occurred and resulted in
	%	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
	%	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
	%	$|\assertionRegion|\neq|\beliefRegion|$. 
	%	\begin{itemize}
	%		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
	%			a communication through the input-port), the two
	%			regions have a different number of subregions, and,
	%			then, there exist only two mutually exclusive options: 
	%			\begin{itemize}
	%				\item either $|\assertionRegion|>|\beliefRegion|$, where one
	%					ore more Assertions didn't become Beliefs, which we
	%					call \emph{drop port} ($s\dropport a$) since its behavior drops and
	%					prevents incoming communications,
	%				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
	%					one ore more Beliefs which have not been sent as
	%					input-Assertions, which we call \emph{insertion
	%					port} ($s\insertport a$) since its behavior generates new Beliefs
	%					unrelated to the incoming communication. 
	%			\end{itemize}
	%		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
	%				there exist only two mutually exclusive options: 
	%			\begin{itemize}
	%				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
	%					$\assertionRegion$ and $\beliefRegion$, i.e.
	%					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
	%					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
	%					$t^0$ and $t^1$ represent time at the beginning of
	%					communication and right after, respectively); which
	%					means that Assertions have reached the port and new
	%					Beliefs have been generated by the port, but no
	%					correlation between the elements of the two regions exist. 
	%					We call this \emph{injection port} ($s\injectionport a$) since it has modified
	%					the information carried by the Assertions into new
	%					corresponding Beliefs.
	%				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
	%					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
	%					this implies that no communication occurred, while we
	%					assumed that a communication was taking place (absurd).
	%			\end{itemize}
	%\end{itemize}
	%In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
	%Similarly, 
	%\begin{itemize}
	%	\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
	%		$\dr{\assertionRegion}{\beliefRegion}$ except for the
	%		overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
	%		$\eq{\assertionRegion}{\beliefRegion}$ (see
	%		\autocite{Santaca2016abf}),
	%		we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
	%	\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
	%		$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
	%		$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
	%	\begin{itemize}
	%		\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
	%			some Beliefs are not correlated to incoming Assertions
	%			and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
	%		\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
	%			some incoming Assertions are not correlated to Belief 
	%			and we call it \emph{selective drop port} ($s\selectivedropport a$)
	%	\end{itemize}
	%\end{itemize}
\end{proof}

\subsubsection{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for the
sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}. A port is, in fact, syntactic sugar to express the 
relation between Assertions and Beliefs.

@@

\paragraph{Channel Weaknesses}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any Assertion from its input-port to its output-port, there is no
insecure behavior but only the combination of the weaknesses of the input and
output port; therefore there exists $(7^2)-2=47$ theoretical configurations
($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on both input and oupput
side; and we exclude the configuration with $2$ secure types as input
and ouput, $-2$); where only 43 are possible.

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{secure output port} and \emph{input injection port} ($s\sjcom r$), 
	\item \emph{secure output port} and \emph{input selective drop port} ($s\ssdcom r$), 
	\item \emph{secure output port} and \emph{input selective insertion port} ($s\ssicom r$), 
	\item \emph{secure output port} and \emph{input selective injection port} ($s\ssjcom r$),

@@

\paragraph{Ports and Channels -- Discussion} 
It is important to note that all the results of the application of the
$\abftheory$ to channels is the analysis of the relation
$\rcc{\assertionRegion_o}{\assertionRegion_i}$ leads to the same results of the
analysis of a pair one input and one output port (i.e.
$\rcc{\beliefRegion_1}{\assertionRegion_2}\wedge\rcc{\assertionRegion_2}{\beliefRegion_3}$
The same result can be obtained by analyzing the relation
between the outputs of a functional block and the inputs of
another functional block.
As depicted in Figure~\ref{fig:atom}, so far, we have considered Information
generated by a Port $P_I$ and then sent through a channel $C$ to another
(recipient) port $P_O$. In this scenario where Ports and then Channels are
atomic (otherwise raising infinite regress), we can only  consider the
relations between Ports and Channel; considering both input-port to channel and
channel to output-port. In fact, the Weaknesses of a channel are defined in terms
of the Weaknesses of Ports. We now introduce the concept of
functional architecture and its decomposition into functional blocks. 
Io order to define a functional block without encountering a infinitely recursive definition,
we must reach the same conclusions as for the Channel. Therefore, describing
the information as flowing over a channel or in a functional block is purely
syntactic sugar.

We can summarize these results by saying that the relations
between Assertions and Beliefs, Output Assertions of an Agent and 
Input Assertions of another Agent, or Output Beliefs of a 
Functional Block and Input Beliefs of another Block can \emph{only}
be affected by the following weaknesses: replace,
drop, injection, insertion, selective drop, and selective drop + insertion.
We can, then, propose the following security properties:
\begin{itemize}
	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
\end{itemize}

@@

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent 
	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
	A functional architecture is an interconnected system of functional blocks.
\end{definition}
It is evident that the quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels). The behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
Therefore, while a port semantics is determined by the relation 
between Assertions and Beliefs, the semantics of a Functional Block 
is determined by the relation between Facts/Requirements and I/O Beliefs.

In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).

\begin{itemize}
	\item $\po{\behaviorRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
	        output given the correct inputs. Not all the inputs are handled
	        properly, nor all the expected outputs are always generated
	        when correct inputs are given.
	\item $\pp{\behaviorRegion}{\factRegion}$ part of the expected outputs are not
	        generated in response to the correct
	        inputs
	\item $\ppi{\behaviorRegion}{\factRegion}$ the components
	        correctly performs the expected behavior when the correct
	        inputs are provided but is subject to input
	        injections
	\item $\dr{\behaviorRegion}{\factRegion}$ the component
		never performs the expected behavior (e.g. physical
		damage)
\end{itemize}

\paragraph{Requirements as Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there may exist a factorial requirement (Fact) predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.3\textwidth]{sum.pdf}
	\caption{Example of a Functional Block}
	\label{fig:sum}
\end{figure}

As an example, consider a functional block that performs the summation of two
inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
+ b_2$. The possible relations between the behavior of the functional block and
the requirements (i.e. $\Rcc{\behaviorRegion}{\factRegion}$ is determined by
the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
$sum(b_1,b_2)=b_1+b_2$, as follows.
\begin{itemize}
	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
		where $\beliefRegion_3$ represents the Region of all the
		outputs of the functional block $sum$ while the Region
		$\beliefRegion_1+\beliefRegion_2$ represents the expected
		outputs of an ideal implementation of the requirement $r$.
		Therefore, the functional block correctly implements the
		requirements.
	\item
		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the function block never implements the requirements.
	\item
		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
		the output is incorrect.
	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
	\item
		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the component has a Byzantine behavior where occasionally
		outputs the expected output given the correct inputs. Not all
		the inputs are handled properly, nor all the expected outputs
		are always generated when correct inputs are given.
\end{itemize}

\paragraph{Methodology}
We have now mapped the building blocks of our theory, namely Information,
Belief, Assertion, Behavior, and Fact. One may notice that the concept of
Knowledge hasn't been discussed in this section so far. Our argument is that
knowledge, as an epistemological concept cannot be defined due to several
implication that this would have (see, for example,
\autocite{Empiricus1990Pyrrhonism}). On the other hand, the factorial way in
which requirements are imposed may be true only in the specification, while not
necessarily true in the final implementation. While Knowledge may not be
apprehensible in general, we can define tests to see if Facts holds in a
design or implementation. What is of our interest is to guarantee security in a
system so we now correlate the potential behaviors with the standard (de-facto)
understanding of security; so that we can identify test that, at the end of the
engineering process, may not give Knowledge but, at last, tested Fact.

@@

\paragraph{Availability and Integrity (Architecture Security)}
\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{client-server.pdf}
	\caption{Client-Server paradigm in $\abftheory$-theory}
	\label{fig:client-server}
\end{figure}
Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
in order to prevent a resource to be available, there
exists only (i.e. inclusively) the following cases
that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
\begin{enumerate}
	\item\label{case:srv1} The resource on the server cannot be found, so that it is
		possible for the client to contact the server and for the server
		to send a response back, but the output-port does not receive the
		necessary information. This may be due to a software error such
		that the functional architecture doesn't output a response, or
		a problem in the flow of information (i.e. the functional
		architecture doesn't receive the necessary information from
		other resources), or an hardware failure \&c. This can be formalized
		as an issue to be found in the relation between the beliefs produced by the functional
		architecture and the ones received by the output port, i.e.
		$\rcc(\behaviorRegion_s,\{\text{res}_0\})$ or 
		$\rcc(\beliefRegion_s,\{\text{res}_0\})$ 
	\item\label{case:srv2} The server hosting the resource is down, and then the server
		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
		requests (insecure input port due to drop-related issue) or both.
	\item\label{case:srv3} The client cannot properly compute request (symmetric to Case~\ref{case:srv1})
	\item The client is down, symmetric to Case~\ref{case:srv2}
	\item The channel is down due to a drop related issue in one of the two channels 
\end{enumerate}
We shall, then, distinguish between two cases, as we mentioned beforehand:
agent (or system) and Information (which we call Infosec) availability. 
\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
when the property of availability is expressed over an agent or a system, the
	list of possible designs do not include those in which there exist:
	drop ports, or drop channels/functional-blocks.
\end{definition}
\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
when the property of availability is expressed
	over a formula, the list of possible designs that violates the
	availability property is reduced to the part of the system that
	does not influence that specific formula.
\end{definition}
